Paper 2: survey analysis

### 2024-09-04 Rmd/Paper2.Rmd
Added correct calculations to reproduce paper 2 results

This document presents the quantitative analyses done in paper 2. Descriptive and inferential analyses are done. The inferences are based on multiple regression analyses.

@@

### presentation of results
#library(pander)                             # prettyprint for tables

@@

@@

The data is multilevel, as the PSTs (level 1) are grouped at the institutional level (level 2).

@@

@@

@@

Should I include interactions in the regression analysis?

@@

```{r TK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

@@

```{r PCK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

@@

```{r TPACK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

@@

## Hierarchical RA for TK

@@

### Level 1

```{r TKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r TKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r TKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r TKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
# interaction
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for PCK

@@

### Level 1

```{r PCKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(PCK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r PCKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(PCK, center=T, scale=T) ~ factor(semester) * (factor(Gender) + factor(Trinn)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

@@

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$PCK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r PCKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r PCKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for TPACK

@@

### Level 1

```{r level1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TPACK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r level1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

@@

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TPACK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

### Level 2

```{r level2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)

#main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r level2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)

fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

### PCK

```{r PCK}
# PCK
m5 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK), 
                clusters = factor(studiested), data=data_clean)
summary(m5)

# main
fstat <- m5$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m5$adj.r.squared / m4$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

### TK

```{r TK}
# TK
m6 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK) + scale(TK), 
                clusters = factor(studiested), data=data_clean)
summary(m6)

fstat <- m6$adj.r.squared / m5$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

>>>>

This document presents the quantitative analyses done in paper 2. Descriptive and inferential analyses are done. The inferences are based on multiple regression analyses. The document is inspired by the idea of a computational essay [@odden_using_2023].

The purpose is foremost to reproduce the results presented in paper 2. But I will also do a reanalysis excluding some interaction effects, which were included in the paper 2 analysis. In addition, I will compare the multiple regression analysis with multilevel modeling, to see if the results are similar, as well as when including cross-level interaction effects due to year of study (semester), since there could be differences in different years either due to changes in the student population or at the teacher education institutes.

@@

### presentation of results
library(pander)                             # prettyprint for tables

@@

Covariates:

* target grade level (Trinn): level 1-7 or level 5-10
* institute: 5 institutes
* gender: male or female
* year of study (semester): 2., 3. or 5. year

@@

The purpose of the multiple regression analysis is to explain the variation in the data. I therefore use stepwise regression and add the variables step-by-step to examine changes in R2 (explained variance).

lm_robust is used to calculate cluster corrected standard errors. (or robust SE? Actually, CR2 SE = cluster robust standard errors)

The data is however multilevel, as the PSTs (level 1) are grouped at the institutional level (level 2). I therefore also have to test whether I need to use hierarchical level modeling (HLM) by examining the intraclass correlation coefficient (ICC) of the data.

@@

ICC is calculated to check whether HLM must be done.

Multiple regression assumes independent observations and linearity between dependent and independent variables. In addition, the residuals should be normally distributed, to calculate test statistics for doing inferences.

@@

The ICC indicates that the data are nested. But since there is only 5 institutes, I used cluster corrected standard errors, instead of HLM, as suggested by @christophersen_introduksjon_2018.

@@

Should I include interactions in the regression analysis? Interaction plots are drawn to see for effects of interaction. Any crossing lines indicate interaction effects. Parallel lines indicate no interaction.

@@

```{r TK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Target grade level",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

@@

```{r PCK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Target grade level",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

@@

```{r TPACK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TPACK,
                 type = "b",
                 ylab = "TPACK",
                 xlab = "Target grade level",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

@@

### Summary of interaction effects

TK level 1 (PST):

* possible interaction between gender and target grade level (Trinn)
* possible interaction between gender and year of study (semester)

TK level 2 (TEI):

* possible interaction between institute and year of study (semester)
* possible weak interaction between institute and target grade level (Trinn)
* possible interaction between gender and institute

PCK level 1:

* possible weak interaction between gender and target grade level
* possible interaction between gender and year of study (semester)
* possible weak interaction between target grade level and year of study

PCK level 2:

* possible interaction between institute and year of study
* possible weak interaction between institute and target grade level 
* possible interaction between institute and gender

TPACK level 1:

* possible interaction between gender and year of study

TPACK level 2:

* possible interaction between institute and year of study
* possible weak interaction between institute and gender

The significance of these interaction effects have to be tested for when doing the regression analysis.

## Replication of paper 2 results

### Hierarchical RA for TK

@@

Interaction effects are examined by calculating F-statistics to see if models with and without interaction are significantly different.

#### Level 1

First, all covariates at the individual PST level 1 (i.e., gender, target grade level and year of study) are introduced as a block.

Without interaction

```{r TKlevel1_main_paper2}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TK_covariates_l1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(model_TK_covariates_l1)
```

With interaction

when calculating interaction effects, TK needs to be standardized [@christophersen_introduksjon_2018].

```{r TKlevel1_interaction_paper2}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TK, center = T, scale = T) ~ factor(Gender) * (factor(Trinn) + factor(semester)),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

Calculate F-stat by comparing R2 to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
df1 <- length(data_clean$TK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

The F-statistic indicate that the two models are not significantly different. This means that interaction effects at TK level 1 can be excluded from the rest of the analyses.

#### Level 2

Then I add level 2-variables: studiested

without interaction

```{r TKlevel2_main_paper2}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_TK_covariates_l2 <- lm_robust(
  scale(TK) ~
    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
    factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_TK_covariates_l2)
```

```{r}
# main
fstat <- model_TK_covariates_l2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- model_TK_covariates_l2$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

with interaction

Trinn is omitted from interaction since it indicated the weakest interaction.

```{r TKlevel2_interaction_paper2}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(
  scale(TK) ~
    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
    factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
  clusters = factor(studiested), data = data_clean
)
summary(m4)
```

Some missing data leads to NA.

```{r}
# interaction
fstat <- m4$adj.r.squared / model_TK_covariates_l2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

The f-statistic indicate no significant differences with or without interaction, so interaction effects at level 2 are excluded from further analysis.

#### Summary

All main variables are included in the further analyses, but no interaction effects.

### Hierarchical RA for PCK

@@

#### Level 1

without interaction, only main effects

```{r PCKlevel1_main_paper2}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_PCK_covariates_l1 <- lm_robust(PCK ~ factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_covariates_l1)
```

with interaction

```{r PCKlevel1_interaction_paper2}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(PCK, center = T, scale = T) ~ 
                  factor(semester) * (factor(Gender) + factor(Trinn)),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

@@

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
df1 <- length(data_clean$PCK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference when including interaction effects. They are therefore excluded from further analysis.

#### Level 2

only main effects

```{r PCKlevel2_main_paper2}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_PCK_covariates_l2 <- lm_robust(
  scale(PCK) ~ 
    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
    factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_covariates_l2)
```

```{r}
# main
fstat <- model_PCK_covariates_l2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- model_PCK_covariates_l2$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference between level 1 and level 2, indicating no difference between TEIs.

interaction effects

target grade level is excluded from the interaction effects

```{r PCKlevel2_interaction_paper2}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(PCK) ~ 
                  factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                  factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / model_PCK_covariates_l2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference when including interaction effects.

#### Summary

Only level 1 variables contribute to explain variance in PCK.

### Hierarchical RA for TPACK

@@

#### Level 1

Covariates are first added as a block.

Main effects

```{r level1_main_paper2}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TPACK_covariates_l1 <- lm_robust(
  TPACK ~
    factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_covariates_l1)
```

Interaction effects

```{r level1_interaction_paper2}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(
  scale(TPACK, center = T, scale = T) ~
    factor(Gender) * (factor(semester) + factor(Trinn)),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

@@

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
df1 <- length(data_clean$TPACK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference when including interaction effects, so they are omitted from further analysis.

#### Level 2

Main effects

```{r level2_main_paper2}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_TPACK_covariates_l2 <- lm_robust(
  scale(TPACK) ~
    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
    factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_covariates_l2)
```

```{r}
#main
fstat <- model_TPACK_covariates_l2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- model_TPACK_covariates_l2$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

Interaction effects

```{r level2_interaction_paper2}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(
  scale(TPACK) ~
    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
    factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
  clusters = factor(studiested), data = data_clean
)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There are no significant interaction effects, so they are excluded from further analysis.

#### PCK

After the covariates are added as a block, PCK is added.

```{r PCK_paper2}
# PCK
model_TPACK_PCK <- lm_robust(
  scale(TPACK) ~
    factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
    factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
    scale(PCK),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_PCK)

# main
fstat <- model_TPACK_PCK$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- model_TPACK_PCK$adj.r.squared / m4$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is a significant difference when adding PCK as a main effect.

#### TK

Finally, TK is added.

```{r TK_paper2}
# TK
model_TPACK_TK <- lm_robust(
  scale(TPACK) ~
    factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
    factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
    scale(PCK) + scale(TK),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_TK)

fstat <- model_TPACK_TK$adj.r.squared / model_TPACK_PCK$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is a significant difference, when adding TK.

### Results summarised

The results differ somewhat from paper 2 in that some interaction effects are excluded, which was included in the analysis done for paper 2. The main trends are however similar.

```{r}
# table

# TK

mtkcl1rs  <- model_TK_covariates_l1$r.squared
mtkcl1ars <- model_TK_covariates_l1$adj.r.squared

mtkcl2rs  <- model_TK_covariates_l2$r.squared
mtkcl2ars <- model_TK_covariates_l2$adj.r.squared

TK_R2     <- data.frame(
  "R2"    = c(mtkcl1rs, mtkcl2rs),
  "R2adj" = c(mtkcl1ars, mtkcl2ars),
  "diff"  = c(mtkcl1ars, mtkcl2ars - mtkcl1ars)
)

rownames(TK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")

pander(TK_R2)
```

```{r}
# PCK

mpcl1rs   <- model_PCK_covariates_l1$r.squared
mpcl1ars  <- model_PCK_covariates_l1$adj.r.squared

mpcl2rs   <- model_PCK_covariates_l2$r.squared
mpcl2ars  <- model_PCK_covariates_l2$adj.r.squared

PCK_R2    <- data.frame(
  "R2"    = c(mpcl1rs, mpcl2rs),
  "R2adj" = c(mpcl1ars, mpcl2ars),
  "diff"  = c(mpcl1ars, mpcl2ars - mpcl1ars)
)

rownames(PCK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute")

pander(PCK_R2)
```

```{r}
# TPACK

mtcl1rs  <- model_TPACK_covariates_l1$r.squared
mtcl1ars <- model_TPACK_covariates_l1$adj.r.squared

mtcl2rs  <- model_TPACK_covariates_l2$r.squared
mtcl2ars <- model_TPACK_covariates_l2$adj.r.squared

mtprs    <- model_TPACK_PCK$r.squared
mtpars   <- model_TPACK_PCK$adj.r.squared

mttrs    <- model_TPACK_TK$r.squared
mttars   <- model_TPACK_TK$adj.r.squared

TPACK_R2 <- data.frame(
  "R2" = c(mtcl1rs, mtcl2rs, mtprs, mttrs),
  "R2adj" = c(mtcl1ars, mtcl2ars, mtpars, mttars),
  "diff" = c(mtcl1ars, mtcl2ars - mtcl1ars, 
             mtpars - mtcl2ars, mttars - mtpars)
)

rownames(TPACK_R2) <- c("Gender, year of study, target grade level", "Teacher education institute", "PCK", "TK")

pander(TPACK_R2)
```

## Reanalysis with interaction effects omitted

### Hierarchical RA for TK

Interaction effects are examined by calculating F-statistics to see if models with and without interaction are significantly different.

#### Level 1

First, all covariates at the individual PST level 1 (i.e., gender, target grade level and year of study) are introduced as a block.

Without interaction

```{r TKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TK_covariates_l1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(model_TK_covariates_l1)
```

With interaction

when calculating interaction effects, TK needs to be standardized [@christophersen_introduksjon_2018].

```{r TKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TK, center = T, scale = T) ~ factor(Gender) * (factor(Trinn) + factor(semester)),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

Calculate F-stat by comparing R2 to see if they are significantly different

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
df1 <- length(data_clean$TK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

The F-statistic indicate that the two models are not significantly different. This means that interaction effects at TK level 1 can be excluded from the rest of the analyses.

#### Level 2

Then I add level 2-variables: studiested

without interaction

```{r TKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_TK_covariates_l2 <- lm_robust(
  scale(TK) ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_TK_covariates_l2)
```

```{r}
# main
fstat <- model_TK_covariates_l2$adj.r.squared / model_TK_covariates_l1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- model_TK_covariates_l2$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

with interaction

Trinn is omitted from interaction since it indicated the weakest interaction.

```{r TKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(
  scale(TK) ~
    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
    factor(studiested) * (factor(semester) + factor(Gender)), # level 2
  clusters = factor(studiested), data = data_clean
)
summary(m4)
```

Some missing data leads to NA.

```{r}
# interaction
fstat <- m4$adj.r.squared / model_TK_covariates_l2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

The f-statistic indicate no significant differences with or without interaction, so interaction effects at level 2 are excluded from further analysis.

#### Summary

All main variables are included in the further analyses, but no interaction effects.

### Hierarchical RA for PCK

#### Level 1

without interaction, only main effects

```{r PCKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_PCK_covariates_l1 <- lm_robust(PCK ~ factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_covariates_l1)
```

with interaction

```{r PCKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(PCK, center = T, scale = T) ~ 
                  factor(Trinn) * (factor(Gender) + factor(semester)) +
                  factor(semester) * factor(Gender),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
df1 <- length(data_clean$PCK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference when including interaction effects. They are therefore excluded from further analysis.

#### Level 2

only main effects

```{r PCKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_PCK_covariates_l2 <- lm_robust(
  scale(PCK) ~ 
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_covariates_l2)
```

```{r}
# main
fstat <- model_PCK_covariates_l2$adj.r.squared / model_PCK_covariates_l1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- model_PCK_covariates_l2$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference between level 1 and level 2, indicating no difference between TEIs.

interaction effects

target grade level is excluded from the interaction effects

```{r PCKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(PCK) ~ 
                  factor(Gender) + factor(Trinn) + factor(semester) +        # level 1
                  factor(studiested) * (factor(semester) + factor(Gender)),  # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / model_PCK_covariates_l2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference when including interaction effects.

#### Summary

Only level 1 variables contribute to explain variance in PCK.

### Hierarchical RA for TPACK

#### Level 1

Covariates are first added as a block.

Main effects

```{r level1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TPACK_covariates_l1 <- lm_robust(
  TPACK ~
    factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_covariates_l1)
```

Interaction effects

```{r level1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(
  scale(TPACK, center = T, scale = T) ~
    factor(Gender) * factor(semester) + factor(Trinn),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
df1 <- length(data_clean$TPACK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference when including interaction effects, so they are omitted from further analysis.

#### Level 2

Main effects

```{r level2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_TPACK_covariates_l2 <- lm_robust(
  scale(TPACK) ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_covariates_l2)
```

```{r}
#main
fstat <- model_TPACK_covariates_l2$adj.r.squared / model_TPACK_covariates_l1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- model_TPACK_covariates_l2$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

Interaction effects

```{r level2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(
  scale(TPACK) ~
    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
    factor(studiested) * (factor(semester) + factor(Gender)), # level 2
  clusters = factor(studiested), data = data_clean
)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / model_TPACK_covariates_l2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

There are no significant interaction effects, so they are excluded from further analysis.

#### PCK

After the covariates are added as a block, PCK is added.
